---
title: "HW4_Group5"
author: "Bhavesh Kilaru"
date: "9/18/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(mice)
library(MASS)
library(forcats)
library(ggplot2)
library(EnvStats)
library(tidyverse)
library(knitr)
```
## loading the data
```{r}
housingData <- read.csv("C:/Users/Bhavesh Kilaru/Desktop/IDA/Home works/HW2/housingData.csv", header = T)
```

## problem 1
#### a
#### creating three new variables in housing data
```{r}
housingData <- housingData %>%
  dplyr::mutate(age = YrSold - YearBuilt,
  ageSinceRemodel = YrSold - YearRemodAdd,
  ageofGarage = YrSold - GarageYrBlt)
```

#### b
#### selecting the numeric columns from housingData data frame
```{r}
#getting the numeric columns
housingNumeric <- select_if(housingData, is.numeric) %>% as_tibble()
head(housingNumeric)
```

#### c
#converting the character columns into factors
```{r}
#housingData[sapply(housingData, is.character)] <- lapply(housingData[sapply(housingData, is.character)], as.factor)
```

#getting the factor columns
```{r}
#housingFactor <-  housingData %>%  select_if(is.factor)
```

#### c
#### creating a tibble housingFactor that contains all factor variables of housing data
```{r}
housingFactor <- housingData %>% transmute(across(where(is.character), as.factor)) %>%as_tibble()
head(housingFactor)
```

#### d
#### using glimse to look at the created tibble
```{r}
glimpse(housingFactor)
```
```{r}
glimpse(housingNumeric)
```

### e
#### functions to extracting Q1 and Q3 
```{r}
Q1<-function(x,na.rm=TRUE) {
quantile(x,na.rm=na.rm)[2]
}
Q3<-function(x,na.rm=TRUE) {
quantile(x,na.rm=na.rm)[4]
}
```

#### f
#### creating the function myNumericSummary
```{r}
myNumericSummary <- function(x){
c(length(x), n_distinct(x), sum(is.na(x)), mean(x, na.rm=TRUE),
min(x,na.rm=TRUE), Q1(x,na.rm=TRUE), median(x,na.rm=TRUE), Q3(x,na.rm=TRUE),
max(x,na.rm=TRUE), sd(x,na.rm=TRUE))
}
```

#### g
#### creating a tibble numericSummary that contains the result of myNumericSummary function
#### applied on each column of housingNumeric tibble
```{r}
numericSummary <- housingNumeric %>% summarise(across(.cols = everything(), ~myNumericSummary(.x)))
glimpse(numericSummary)
```

#### h
#### adding a column stats that names each row of numericSummary
```{r}
numericSummary <-cbind( stat=c("n","unique","missing","mean","min","Q1","median","Q3","max","sd"),
                        numericSummary)
glimpse(numericSummary)
```
#### i
#### pivoting the data
```{r}
numericSummaryFinal <- numericSummary %>%
  pivot_longer("Id":"ageofGarage", names_to = "variable", values_to = "value") %>%
  pivot_wider(names_from = stat, values_from = value) %>%
  mutate(missing_pct = 100*missing/n,
         unique_pct = 100*unique/n) %>%
  dplyr::select(variable, n, missing, missing_pct, unique, unique_pct, everything())

numericSummaryFinal
```
```{r}
options(digits=3)
options(scipen=99)
numericSummaryFinal %>% kable()
```

##### j
##### getting the modes
```{r}
getmodes <- function(v,type=1) {
  tbl <- table(v)
  m1<-which.max(tbl)
  if (type==1) {
    return (names(m1)) #1st mode
  }
  else if (type==2) {
    return (names(which.max(tbl[-m1]))) #2nd mode
  }
  else if (type==-1) {
    return (names(which.min(tbl))) #least common mode
  }
  else {
    stop("Invalid type selected")
  }
}
```

#### getting mode cnts
```{r}
getmodesCnt <- function(v,type=1) {
  tbl <- table(v)
  m1<-which.max(tbl)
  if (type==1) {
    return (max(tbl)) #1st mode freq
  }
  else if (type==2) {
    return (max(tbl[-m1])) #2nd mode freq
  }
  else if (type==-1) {
    return (min(tbl)) #least common freq
  }
  else {
    stop("Invalid type selected")
  }
}
```

##### creating the MyFactorSummary function to summarize factor columns
```{r}
myFactorSummary<-function(x){
  c(length(x), n_distinct(x), sum(is.na(x)), 
    getmodes(x, type =1), getmodesCnt(x, type =1),
    getmodes(x, type =2), getmodesCnt(x, type =2),
    getmodes(x, type =-1), getmodes(x, type =-1))
}
```

#### creating a tibble FactorSummary that contains the result of myFactorSummary function
#### applied on each column of housingFactor tibble
```{r}
FactorSummary <- housingFactor %>% summarise(across(.cols = everything(), ~myFactorSummary(.x)))
glimpse(FactorSummary)
```
#### adding a column stats that names each row of FactorSummary
```{r}
FactorSummary <-cbind( stat=c("n","unique","missing","1st Mode", "1st Mode Freq", "2nd Mode", 
                              "2nd Mode Freq", "Least Common Mode", "Least Common Mode Freq"),
                        FactorSummary)
glimpse(FactorSummary)
```

#### pivoting the data
```{r}
FactorSummaryFinal <- FactorSummary %>%
  pivot_longer("MSZoning":"SaleType", names_to = "variable", values_to = "value") %>%
  pivot_wider(names_from = stat, values_from = value) %>%
  mutate(missing_pct = 100* as.numeric(missing)/as.numeric(n),
         unique_pct = 100* as.numeric(unique)/as.numeric(n)) %>%
  dplyr::select(variable, n, missing, missing_pct, unique, unique_pct, everything())

FactorSummaryFinal
```

#### converting data to kable
```{r}
options(digits=3)
options(scipen=99)
FactorSummaryFinal %>% kable()
```

## 2
#### a
#### plotting the histogram of all numeric columns in original data to check the skewness
```{r}
par(mar = c(0.75, 0.75, 0.75, 0.75))
par(mfrow=c(10,4))
housingNumeric <- select_if(housingData, is.numeric)
for(i in 1:ncol(housingNumeric)){
  hist(housingNumeric[, i], main = colnames(housingNumeric)[i], axes = F)
}
```

```{r}
for (i in 1:ncol(housingNumeric)){
  print(paste(colnames(housingNumeric)[i], skewness(housingNumeric[, i])))
}
```

```{r}
#boxcox(housingNumeric$OpenPorchSF, optimize=T, lambda=c(-3,3))
```

##### Here I consider columns PoolArea, LotArea and BsmtUnfSFas they are continuous variables

##### Pool Area
#####taking pre snap of PoolArea attribute
```{r}
temp_PoolArea <- housingNumeric$PoolArea
```

##### box cox of PoolArea Attribute
##### adding a negligible value to PoolArea since boxcox is infeasible when value is zero
```{r}
b <- boxcox(housingNumeric$PoolArea+0.0001, optimize = T, lambda = c(-3, 3))
#changing the values of PoolArea using boxcox function
housingNumeric$PoolArea <- (housingNumeric$PoolArea^b$lambda - 1)/b$lambda
```

##### post snap of PoolArea
```{r}
par(mfrow = c(2, 1))

#plotting a pre snap of PoolArea attribute
hist(temp_PoolArea, xlab = 'PoolArea', main ='Presnap of PoolArea')

#plotting the post snap of PoolArea
hist(housingNumeric$PoolArea, xlab = 'PoolArea', main ='Post snap of PoolArea')
```
since most of the data is concentrated at zero, PoolArea haven't changed even after applying the
BoxCox function


#### Lot AREA
#####getting pre snap of Lot area
```{r}
temp_LotArea <- housingNumeric$LotArea
```
##### boxcox of Lot area
```{r}
b <- boxcox(housingNumeric$LotArea, optimize = T, lambda = c(-3,3))

#changing the values of LotArea using boxcox function
housingNumeric$LotArea <- (housingNumeric$LotArea^b$lambda -1)/b$lambda
```
##### plotting pre snap and post snap of Lot area
```{r}
par(mfrow=c(2, 1))

#plotting the pre snap
hist(temp_LotArea, xlab = 'Lot Area', main ='Presnap of Lot Area')

#plotting the post snap
hist(housingNumeric$LotArea, xlab = 'Lot Area', main ='postsnap of Lot Area')
```

#### BsmtUnfSF

##### taking a pre snap of BsmtUnfSF
```{r}
temp_BsmtUnfSF <- housingNumeric$BsmtUnfSF
```

##### box cox of BsmtUnfSF Attribute
##### adding a negligible value to BsmtUnfSF since boxcox is infeasible when value is zero
```{r}
b <- boxcox(housingNumeric$BsmtUnfSF+0.0001, optimize = T, lambda = c(-3, 3))
#changing the values of MSsubClass using boxcox function
housingNumeric$BsmtUnfSF <- (housingNumeric$BsmtUnfSF^b$lambda - 1)/b$lambda
```

##### post snap of MSSub Class
```{r}
par(mfrow = c(2, 1))

#plotting a pre snap of BsmtUnfSF attribute
hist(temp_BsmtUnfSF, xlab = 'BsmtUnfSF', main ='Presnap of BsmtUnfSF')

#plotting the post snap of BsmtUnfSF
hist(housingNumeric$BsmtUnfSF, xlab = 'BsmtUnfSF', main ='Post snap of BsmtUnfSF')
```

##### b

##### mean imputation of missing values in LotFrontage Column
```{r}
#creating a temporary variable to store the original housing data
data_imp_mean <- housingData

#getting pre count
print(paste("The number of missing values before mean impuatation is ", sum(is.na(data_imp_mean$LotFrontage))))

#getting the position of missing values of LotFrontage attribute
missing_values <- which(is.na(data_imp_mean$LotFrontage))

#imputing missing values of LotFrontage attribute with mean
data_imp_mean$LotFrontage[missing_values] <- mean(data_imp_mean$LotFrontage, na.rm = T)

#getting post count
print(paste("The number of missing values after mean impuatation is ", sum(is.na(data_imp_mean$LotFrontage))))
```

##### ii
##### regression with error
```{r}
#getting a copy of original housing data
data_imp_RE <- housingData

#getting pre count
print(paste("The number of missing values before mean impuatation is ", sum(is.na(data_imp_RE$LotFrontage))))
```
```{r}
#fitting the linear regression model
fit<-lm(LotFrontage ~ LotArea + SalePrice, data_imp_RE)  

#getting the summary of the model
f<-summary(fit)
print (f)  
```
```{r}
# extract the coefficients 
c<-f[[4]]      

# extract the model standard error
se<-f[[6]]                    
```

```{r}
#getting the position of missing values of LotFrontage attribute
missing <- which(is.na(data_imp_RE$LotFrontage))

#data_imp_RE[missing,"LotFrontage"]<- (c[1] + c[2]*data_imp_RE[missing,"LotArea"] + c[3] * data_imp_RE[missing,"SalePrice"] )

#predicting missing values using the linear model fitted and adding error to it
data_imp_RE[missing,"LotFrontage"]<- predict(fit, housingData[missing, c('LotArea', 'SalePrice')]) +  rnorm(length(missing),0,se)
```

```{r}
#getting post count
print(paste("The number of missing values after mean impuatation is ", sum(is.na(data_imp_RE$LotFrontage))))
```

##### iii
#####  predictive mean matching
```{r}
#getting a copy of original housing data
temp_HD <- read.csv("C:/Users/Bhavesh Kilaru/Desktop/IDA/Home works/HW2/housingData.csv", header = T)

#getting pre count
print(paste("The number of missing values before mean impuatation is ", sum(is.na(temp_HD$LotFrontage))))

# Impute missing values using PMM
imp_single <- mice(temp_HD, m = 1, method = "pmm")

#getting only the complete cases from above 'imp_single' mids class
data_imp_pmm <- complete(imp_single)

#getting post count
print(paste("The number of missing values after mean impuatation is ", sum(is.na(data_imp_pmm$LotFrontage))))
```

##### Histogram of Lot Frontage after imputation using predictive mean matching
```{r}
#pre snap of lot frontage
hist(housingData$LotFrontage,xlab = 'LotFrontage', 
     main = "pre snap of LotFrontage Column")
```

```{r}
#post snap of LotFrontage after after mean imputation
hist(data_imp_mean$LotFrontage, xlab = "Lot Frontage",
     main = "Histogram of Lot Frontage after mean imputation")
```

```{r}
#post snap of LotFrontage after imputation using regression with error
hist(data_imp_RE$LotFrontage, xlab = 'Lot Frontage', 
     main = 'Histogram of Lot Frontage after imputation using regression with error')
```

```{r}
#post snap of LotFrontage after imputation using PMM
hist(data_imp_pmm$LotFrontage, xlab = 'Lot Frontage', 
     main = 'Histogram of Lot Frontage after imputation using pmm')
```

##### c
##### getting the pre count of each factor in attribute Exterior1st 
#####in descending order
```{r}
#summary(housingData$Exterior1st)
fct_count(housingData$Exterior1st, sort = T)
```

##### collapsing the columns so that the column Exterior1st have only five factors
```{r}
housingData$Exterior1st <- fct_collapse(housingData$Exterior1st, 
                                          Other = c("BrkFace", "CemntBd", "other", "Plywood"))

#getting the post count for each factor level in descending order
fct_count(housingData$Exterior1st, sort = T)
```

#### d
##### i
##### calculating average sales price for each neighborhood
```{r}
housingData %>% group_by(Neighborhood) %>% 
  summarise(Mean = mean(SalePrice))
```

##### ii
##### making the boxplots of Saleprice for each neighborhood
```{r}
ggplot(housingData, aes(x= Neighborhood, y = SalePrice)) +
  #Adding the boxplot layer
  geom_boxplot() + 
  #adding title
  labs(title = "box plot for each of the neighborhood", xlab = 'Neighborhood') +
  #setting the title of the plot to middle and alligning the xlabs vertically
  theme(axis.text.x = element_text(angle = 90), plot.title = element_text(hjust = 0.5))
```

##### iii
##### using forcats, ordering the factors levels of neighborhooding based on the decsending order of Median of Saleprice
```{r}
housingData$Neighborhood <-fct_reorder(housingData$Neighborhood, 
                                       housingData$SalePrice, .fun = median, .desc = T)
```

##### displaying the first five rows based on above transformation
```{r}
head(housingData %>% group_by(Neighborhood) %>% 
  summarise(Median = median(SalePrice)))
```

##### iv
#### making boxplot of saleprice for each neighborhood which is sorted on descending order of median
```{r}
ggplot(housingData, aes(x= Neighborhood, y = SalePrice)) +
  #Adding the boxplot layer
  geom_boxplot() + 
  #adding title
  labs(title = "box plot for each of the neighborhood after reordering", xlab = 'Neighborhood') +
  #setting the title of the plot to middle and alligning the xlabs vertically
  theme(axis.text.x = element_text(angle = 90), plot.title = element_text(hjust = 0.5))
```





